---
title: "Examining the citation and altmetric advantage of bioRxiv preprints: bioRxiv figures"
---

### Initial Configuration

```{r results='hide'}

# Call libraries
library(tidyverse)
library(lubridate)
library(ggpubr)
library(viridis) # better color palettes
library(eulerr) # For Venn diagrams


# Theme for plots
new_theme <- theme_bw() +
  theme(axis.title.x.bottom=element_text(margin=margin(t=10, r=0, b=0, l=0)),
        axis.title.y.left=element_text(margin=margin(t=0, r=10, b=0, l=0)),
        axis.title.x.top=element_text(margin=margin(t=1, r=0, b=10, l=0)),
        axis.title.y.right=element_text(margin=margin(t=0, r=0, b=, l=10)),
        axis.text=element_text(size=10, colour = "grey30"),
        axis.title=element_text(size=10),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        panel.border = element_rect(colour = "grey30"),
        legend.title = element_blank(),
        strip.background = element_blank(),
        strip.text.x=element_text(size=10, colour = "grey30"))

theme_set(new_theme)

# Figure variables
width <- 20
height_small <- 8
height_med <- 12
height_large <- 16
resolution <- 900
units <- "cm"

```

### Figure 1: Overlap in preprint-published article links

```{r results='hide'}

png("outputs/figures/fig_1_matching_overlap.png", 
    width = width, 
    height = height_small, 
    units=units, 
    res=resolution)

# Load data
fit <- deframe(read_csv("data/analysis/matching_overlap.csv"))

# Plot Venn diagram
#plot(euler(fit), 
#     labels=c("Crossref", "bioRxiv", "Scopus"), 
#     quantities=TRUE, 
#     fill = c("#993366", "#2166AC", "#E69F00"),
#     alpha = 0.5,
#     cex = 2)

plot(venn(fit), 
     labels=c("Crossref", "bioRxiv", "Scopus"), 
     fill = c("#993366", "#2166AC", "#E69F00"),
     color = '#000000',
     alpha = 0.5)

dev.off()

rm(fit)

```

### Figure 2: Monthly evolution of bioRxiv submissions and publications

```{r results='hide'}

# Number of preprints deposited over time
p1 <- read_csv("data/analysis/monthly_preprints.csv") %>%
  ggplot(aes(x=YEAR_MON, y=DEPOSITED)) +
  geom_bar(stat="identity", color="black", fill="grey", lwd=0.25) + 
  labs(x = "", y="Preprints\nDeposited") +
  theme(panel.spacing.x = unit(0, "cm"), # plot facet panels together
        axis.text.x = element_blank()) + # hide months from x-axis)
  facet_grid(~YEAR, scales="free", space="free_x", switch="both")

# Percentage of preprints published over time
p2 <- read_csv("data/analysis/monthly_preprints.csv") %>%
  ggplot(aes(x=YEAR_MON, y=PCT_PUBLISHED)) +
  geom_bar(stat="identity", color="black", fill="grey", lwd=0.25) + 
  labs(x = "", y="Percentage\nPublished") +
  lims(y=c(0,100)) +
  theme(panel.spacing.x = unit(0, "cm"), # plot facet panels together
        axis.text.x = element_blank()) +
  facet_grid(~YEAR, scales="free", space="free_x", switch="both")

# Arrange plot panels
plot <- ggarrange(p1, p2, labels = c("A", "B"), ncol = 1, nrow = 2, align="v")

# Save plot to .png
ggsave("outputs/figures/fig_2_monthly_preprints.png", 
       plot, 
       width=width, 
       height=height_med, 
       units=units, 
       dpi=resolution)

# Remove redundant variables
rm(p1, p2, plot)

```

### Figure 3: Monthly citation rates of bioRxiv-deposited and control articles

```{r}

d <- read_csv("data/analysis/article_monthly_citations.csv") %>%
  mutate(TYPE = factor(TYPE, levels=c("Deposited", "Control")))

p1 <- d %>%
  group_by(TYPE, CITATION_INTERVAL) %>%
  summarize(CPP_MEAN = mean(LOG_CITATIONS),
            CPP_95CI = qnorm(0.975)*sd(LOG_CITATIONS)/sqrt(n())) %>%
  ggplot(aes(x=CITATION_INTERVAL, group=TYPE)) +
  geom_ribbon(aes(ymin=CPP_MEAN-CPP_95CI, ymax=CPP_MEAN+CPP_95CI), fill="grey", alpha=0.3) +
  geom_line(aes(y=CPP_MEAN, color=TYPE)) +
  geom_point(aes(y=CPP_MEAN, fill=TYPE, color=TYPE)) +
  scale_fill_manual(values=c("#2166AC", "#E69F00")) +
  scale_color_manual(values=c("#2166AC", "#E69F00")) +
  coord_cartesian(xlim=c(0,36), ylim=c(0,0.6)) +
  scale_x_continuous(breaks=c(0,6,12,18,24,30,36), position="top") +
  xlab("Months after preprint deposit") +
  ylab(expression(italic("Cpp"))) +
  theme(legend.position = c(0.15, 0.85), # manual shuffling
        legend.background = element_rect(fill="white", 
                                         size=0.2, 
                                         color="grey30", 
                                         linetype="solid"),
        legend.direction = "horizontal",
        plot.margin = margin(0,0,0.75,1,"cm"))

p2 <- d %>%
  group_by(PREPRINT_DOI, CITATION_INTERVAL) %>%
  arrange(TYPE) %>%
  summarize(PAIRED_DIFFERENCE = (LOG_CITATIONS[1]-LOG_CITATIONS[2])) %>%
  ungroup() %>%
  group_by(CITATION_INTERVAL) %>%
  summarize(DIFF_MEAN = mean(PAIRED_DIFFERENCE),
            DIFF_95CI = qnorm(0.975)*sd(PAIRED_DIFFERENCE)/sqrt(n())) %>%
  ungroup() %>%
  ggplot(aes(x=as.numeric(CITATION_INTERVAL))) +
  geom_line(aes(y=DIFF_MEAN)) +
  geom_point(aes(y=DIFF_MEAN)) +
  geom_ribbon(aes(ymin=DIFF_MEAN-DIFF_95CI, ymax=DIFF_MEAN+DIFF_95CI), alpha=0.1) +
  geom_hline(yintercept = 0, linetype="dashed", colour="black", lwd=0.25) +
  coord_cartesian(xlim=c(0,36), ylim=c(-0.3, 0.3)) +
  scale_x_continuous(breaks=c(0,6,12,18,24,30,36)) +
  xlab("") +
  ylab(expression(Delta~italic("Cpp")["paired"])) +
  theme(axis.text.x = element_blank(),
        plot.margin = margin(0,0,0,1,"cm"))

p3 <- d %>%
  filter(TYPE == "Deposited") %>%
  group_by(CITATION_INTERVAL) %>%
  summarize(n = n()) %>%
  ggplot(aes(x=CITATION_INTERVAL, y=n)) +
  geom_bar(stat="identity", fill="grey", color="black", lwd=0.25) +
  coord_cartesian(xlim=c(0,36)) +
  scale_x_continuous(breaks=c(0,6,12,18,24,30,36)) +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000)) +
  xlab("Months after publication") +
  ylab("Sample size") +
  theme(panel.grid.minor = element_blank(),
        plot.margin = margin(0,0,0,1,"cm"))

plot <- ggarrange(p1, p2, p3, labels = c("A", "B", "C"),
                  ncol = 1, nrow = 3, align="v", heights=c(1.8,1.5,1)) #manual shuffling

ggsave("outputs/figures/fig_3_monthly_articles_citations.png", 
       plot,
       width=width, 
       height=height_large, 
       units="cm", 
       dpi=resolution)

```

### Figure 4: Monthly citation rates for published and unpublished preprints

```{r}

# Read data
d <- read_csv("data/analysis/all_preprints_monthly_citations.csv") 

# Monthly citations rates of published and unpublished preprints
p1 <- d %>%
  group_by(IS_PUBLISHED, CITATION_INTERVAL) %>%
  summarize(
    CPP_MEAN = mean(LOG_CITATIONS),
    CPP_95CI = qnorm(0.975)*sd(LOG_CITATIONS)/sqrt(n())
  ) %>% 
  ggplot(aes(x=CITATION_INTERVAL, y=CPP_MEAN, group=IS_PUBLISHED)) +
  geom_ribbon(aes(ymin=CPP_MEAN-CPP_95CI, ymax=CPP_MEAN+CPP_95CI), linetype=2, alpha=0.1) +
  geom_line(aes(colour=IS_PUBLISHED), size=0.5) +
  geom_point(aes(colour=IS_PUBLISHED), shape=21, size=1, stroke=0.75, fill="white")  +
  scale_x_continuous(limits=c(0, 36), 
                     breaks = c(0, 6, 12, 18, 24), 
                     position="top") +
  coord_cartesian(ylim=c(0,0.03), xlim=c(0,24)) +
  scale_color_manual(values=c("#2166AC", "#E69F00")) +
  xlab("Months after preprint deposit") +
  ylab(expression(italic("Cpp"))) +
  theme(legend.position = c(0.18, 0.9),
        legend.background = element_rect(fill="white", 
                                         size=0.2, 
                                         color="grey30", 
                                         linetype="solid"),
        legend.direction = "horizontal")
  
# Sample size for each group at each time interval
p2 <- d %>%
  group_by(IS_PUBLISHED, CITATION_INTERVAL) %>%
  summarize(
    N = n()
  ) %>% 
  ggplot(aes(x=CITATION_INTERVAL, y=N)) +
  geom_bar(aes(fill=IS_PUBLISHED), stat="identity") +
  scale_x_continuous(limits=c(-1, 25), breaks = c(0, 6, 12, 18, 24)) +
  scale_y_continuous(breaks = c(0, 4000, 8000, 12000, 16000, 20000)) +
  scale_fill_manual(values=c("#2166AC", "#E69F00")) +
  coord_cartesian(xlim=c(0,24)) +
  xlab("Months after preprint deposit") +
  ylab("Sample size") +
  theme(plot.margin = margin(0, 0, 0, 1, "cm"),
        panel.grid.minor = element_blank(),
        legend.position = c(0.82, 0.75),
        legend.background = element_rect(fill="white", 
                                         size=0.2, 
                                         color="grey30", 
                                         linetype="solid"),
        legend.direction = "horizontal")

# Arrange panels
plot <- ggarrange(p1, p2, labels = c("A", "B"), 
                  ncol = 1, nrow = 2, align="v", heights=c(2,1))

# Save plot
ggsave("outputs/figures/fig_4_monthly_preprints_cpp.png", 
       plot, 
       width=width, 
       height=height_med,
       units="cm",
       dpi=resolution)

rm(p1, p2, plot)

```

### Figure 5: Monthly citation rates for preprints as a function of time before/after journal publication

```{r}

# Read data
d <- read_csv("data/analysis/published_preprints_monthly_citations.csv") 

# Monthly citation rates for preprints as a function of journal article publication data
p1 <- d %>%
  group_by(CITATION_INTERVAL) %>%
  summarize(
    CPP_MEAN = mean(LOG_CITATIONS),
    CPP_95CI = qnorm(0.975)*sd(LOG_CITATIONS)/sqrt(n())
  ) %>% 
  ggplot(aes(x=CITATION_INTERVAL, y=CPP_MEAN)) +
  geom_ribbon(aes(ymin=CPP_MEAN-CPP_95CI, ymax=CPP_MEAN+CPP_95CI), linetype=2, alpha=0.1) +
  geom_line(size=0.5) +
  geom_point(shape=21, size=1, stroke=0.75, fill="white") +
  geom_vline(xintercept=0, linetype="dotted") +
  geom_text(x=0, y=0,label="Journal article\npublished", size=3, angle = 90, hjust = 0) +
  xlab("Months relative to journal article publication") +
  ylab(expression(italic("Cpp"))) +
  scale_x_continuous(limits=c(-12, 24), breaks = c(-12, -6, 0, 6, 12, 18, 24), position="top") +
  coord_cartesian(ylim=c(0,0.03))
  
# Sample size at each time interval
p2 <- d %>%
  group_by(CITATION_INTERVAL) %>%
  summarize(
    N = n()
  ) %>% 
  ggplot(aes(x=CITATION_INTERVAL, y=N)) +
  geom_bar(stat="identity") +
  scale_x_continuous(limits=c(-13, 25), breaks = c(-12, -6, 0, 6, 12, 18, 24)) +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000)) +
  coord_cartesian(xlim=c(-12, 24)) +
  xlab("Months relative to journal article publication") +
  ylab("Sample size") +
  theme(plot.margin = margin(0, 0, 0, 1, "cm"),
        panel.grid.minor = element_blank())

# Arrange panels
plot <- ggarrange(p1, p2, labels = c("A", "B"), 
                  ncol = 1, nrow = 2, align="v", heights=c(2,1))

# Save plot
ggsave("outputs/figures/fig_5_monthly_published_preprints_cpp.png", 
       plot, 
       width=width, 
       height=height_med, 
       units="cm", 
       dpi=resolution)

rm(p1, p2, plot)

```

### Figure 6: Altmetrics of preprints, bioRxiv-deposited and control articles

```{r results='hide'}

preprint_altmetrics <- read_csv("data/raw/preprint_altmetrics.csv") %>%
  mutate(TYPE="Preprint") 

article_altmetrics <- read_csv("data/raw/article_altmetrics.csv") %>%
  inner_join(select(read_csv("data/analysis/article_characteristics.csv"), 
                    PREPRINT_DOI, ARTICLE_DOI, TYPE), 
             by=c("PREPRINT_DOI", "ARTICLE_DOI"))

d <- bind_rows(preprint_altmetrics, article_altmetrics) %>%
  mutate(TYPE = factor(TYPE, levels=c("Preprint", "Deposited", "Control")))

# Altmetrics coverage
p1 <- d %>%
  group_by(TYPE) %>%
  summarize(PCT_HAS_ALTMETRIC = (sum(HAS_ALTMETRIC)/n())*100) %>%
  ggplot(aes(x=TYPE, y=PCT_HAS_ALTMETRIC, fill=TYPE)) +
  geom_bar(colour="black", stat="identity", position=position_dodge()) +
  xlab("Altmetrics Coverage") +
  ylab("%") +
  scale_fill_manual(values=c("#993366", "#2166AC", "#E69F00")) +
  theme(axis.text.x = element_blank())

# Mean tweets
p2 <- d %>%
  group_by(TYPE) %>%
  summarize(TWEETS_MEAN = mean(log(ALT_TWEETS+1)),
            TWEETS_95CI = qnorm(0.975)*sd(log(ALT_TWEETS+1))/sqrt(n())) %>%
  ggplot(aes(x=TYPE, y=TWEETS_MEAN, fill=TYPE)) +
  geom_bar(colour="black", stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=TWEETS_MEAN-TWEETS_95CI, ymax=TWEETS_MEAN+TWEETS_95CI), 
                width=.2, position=position_dodge(.9)) +
  xlab("Twitter") +
  ylab("Mean Log-Counts") +
  scale_fill_manual(values=c("#993366", "#2166AC", "#E69F00")) +
  theme(axis.text.x = element_blank())

# Mean blog  mentions
p3 <- d %>%
  group_by(TYPE) %>%
  summarize(FEEDS_MEAN = mean(log(ALT_FEEDS+1)),
            FEEDS_95CI = qnorm(0.975)*sd(log(ALT_FEEDS+1))/sqrt(n())) %>%
  ggplot(aes(x=TYPE, y=FEEDS_MEAN, fill=TYPE)) +
  geom_bar(colour="black", stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=FEEDS_MEAN-FEEDS_95CI, ymax=FEEDS_MEAN+FEEDS_95CI), 
                width=.2, position=position_dodge(.9)) +
  xlab("Blogs") +
  ylab("Mean Log-Counts") +
  scale_fill_manual(values=c("#993366", "#2166AC", "#E69F00")) +
  theme(axis.text.x = element_blank())

# Mean mainstream media mentions
p4 <- d %>%
  group_by(TYPE) %>%
  summarize(MSM_MEAN = mean(log(ALT_MSM+1)),
            MSM_95CI = qnorm(0.975)*sd(log(ALT_MSM+1))/sqrt(n())) %>%
  ggplot(aes(x=TYPE, y=MSM_MEAN, fill=TYPE)) +
  geom_bar(colour="black", stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=MSM_MEAN-MSM_95CI, ymax=MSM_MEAN+MSM_95CI), width=.2, position=position_dodge(.9)) +
  xlab("News") +
  ylab("Mean Log-Counts") +
  scale_fill_manual(values=c("#993366", "#2166AC", "#E69F00")) +
  theme(axis.text.x = element_blank())

# Mean wikipedia mentions
p5 <- d %>%
  group_by(TYPE) %>%
  summarize(WIKIPEDIA_MEAN = mean(log(ALT_WIKIPEDIA+1)),
            WIKIPEDIA_95CI = qnorm(0.975)*sd(log(ALT_WIKIPEDIA+1))/sqrt(n())) %>%
  ggplot(aes(x=TYPE, y=WIKIPEDIA_MEAN, fill=TYPE)) +
  geom_bar(colour="black", stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=WIKIPEDIA_MEAN-WIKIPEDIA_95CI, ymax=WIKIPEDIA_MEAN+WIKIPEDIA_95CI), width=.2, position=position_dodge(.9)) +
  xlab("Wikipedia") +
  ylab("Mean Log-Counts") +
  scale_fill_manual(values=c("#993366", "#2166AC", "#E69F00")) +
  theme(axis.text.x = element_blank())

# Mean Mendeley reads
p6 <- d %>%
  group_by(TYPE) %>%
  summarize(MENDELEY_MEAN = mean(log(ALT_MENDELEY+1)),
            MENDELEY_95CI = qnorm(0.975)*sd(log(ALT_MENDELEY+1))/sqrt(n())) %>%
  ggplot(aes(x=TYPE, y=MENDELEY_MEAN, fill=TYPE)) +
  geom_bar(colour="black", stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=MENDELEY_MEAN-MENDELEY_95CI, ymax=MENDELEY_MEAN+MENDELEY_95CI), width=.2, position=position_dodge(.9)) +
  xlab("Mendeley") +
  ylab("Mean Log-Counts") +
  scale_fill_manual(values=c("#993366", "#2166AC", "#E69F00")) +
  theme(axis.text.x = element_blank())

# Paired altmetric indicators
p7 <- d %>% filter(TYPE %in% c("Deposited", "Control")) %>%
  select(PREPRINT_DOI, TYPE, ALT_TWEETS, ALT_FEEDS, ALT_MSM, ALT_WIKIPEDIA, ALT_MENDELEY) %>%
  gather(key="INDICATOR", value="VALUE", -PREPRINT_DOI, -TYPE) %>%
  group_by(PREPRINT_DOI, INDICATOR) %>%
  arrange(TYPE) %>%
  mutate(PAIRED_DIFFERENCE = log(VALUE[1]+1)-log(VALUE[2]+1)) %>%
  ungroup() %>%
  group_by(INDICATOR) %>%
  summarize(DIFF_MEAN = mean(PAIRED_DIFFERENCE),
            DIFF_95CI = qnorm(0.975)*sd(PAIRED_DIFFERENCE)/sqrt(n())) %>%
  mutate(INDICATOR =case_when(
    INDICATOR == "ALT_FEEDS" ~ "Blogs",
    INDICATOR == "ALT_MENDELEY" ~ "Mendeley",
    INDICATOR == "ALT_TWEETS" ~ "Twitter",
    INDICATOR == "ALT_MSM" ~ "News",
    INDICATOR == "ALT_WIKIPEDIA" ~ "Wikipedia"
  )) %>%
  mutate(INDICATOR = factor(INDICATOR, levels=c("Twitter", "Blogs", "News", "Wikipedia", "Mendeley"))) %>%
  ggplot(aes(x=INDICATOR, y=DIFF_MEAN)) +
  geom_bar(colour="black", fill="grey", stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=DIFF_MEAN-DIFF_95CI, ymax=DIFF_MEAN+DIFF_95CI), width=.2, position=position_dodge(.9)) +
  geom_hline(yintercept=0, color="black", linetype="dashed", lwd=0.25) +
  coord_cartesian(ylim=c(-1,1)) +
  theme(axis.text.x = element_text(angle = 90)) +
  xlab("") +
  ylab(expression(Delta~italic("Log-Counts")["paired"])) +
  theme(plot.margin = margin(2, 1, 0, 1, "cm"))
  
# Arrange panels
plot <- ggarrange(p1, p2, p3, p4, p5, p6, ncol = 3, nrow = 2, align="hv", common.legend = TRUE, legend="top")

plot_full <- ggarrange(plot, p7, labels=c("A", "B"), ncol=2, nrow=1, widths=c(1.5,1))

# Save plot
ggsave("outputs/figures/fig_6_altmetrics_summary.png", 
       plot_full,
       width=width*1.5, 
       height=height_med, 
       units="cm", 
       dpi=resolution)

rm(p1, p2, p3, p4, p5, p6, p7, plot, article_altmetrics)

```
